% Document class
\documentclass[11pt]{article}

% Import packages
\input{packages}
\usepackage{custom_macros}

\pagestyle{fancy}

% Can you suppress the first running heading?

\begin{document}
\vspace{-2cm}
\begin{center}
    % Subject name
    {\Large\bf Torque of Concentric Elliptic Rings} \\
    \vspace{5mm}
    %\vspace{-3cm}
    {\Large Pavadol Yamsiri} \\
    \vspace{3mm}
    Supervisors: Joss Bland-Hawthorn, Thor Tepper-Garcia \\
    \vspace{2mm}
    Sydney Institute for Astronomy
\end{center}
\vspace{0.5mm}

% TODO: Add bibtex references
\begin{abstract}
    % find earliest ref. to this - probably first ever discussion of spiral arms or soon after.
    Astronomers have long known that they can produce spiral perturbations in a disc by using concentric elliptic annuli
    (or rings) that are rotated as a slow function of the ellipse's major axis position angle (Lindblad 1956). To our knowledge,
    this idea has only ever been explored geometrically (e.g. Kalnajs 1973) and never used as a toy model involving kinematic or
    dynamic arguments. Inspired by recent N-body simulations of an impulse-driven density wave in a cold stellar disc, we
    investigate the nature of that wave. With the aid of a fully dynamical toy model in 2D and 3D, we see that the early spiral
    wave (seen to wrap up as time passes) is entirely kinematic, with no forces acting. The initial impulse perturbs the disc
    such that the low-order vertical and radial frequencies are simple perturbations of the disc potential. In the plane, this
    gives rise to coherent rotation of the ellipses (see above) following Lindblad's formula \( \Omega(R)-\kappa(R)/2 \) where
    \( \Omega \) is the angular frequency of an ellipse's major axis, and \( \kappa \) is the radial frequency. But at late times,
    the wrapping stalls and the kinematic wave loses its coherence. At this point, the spiral arms become a dynamical density
    wave, where the inter-arm forces lead to instability, as seen in the simulations. We see this behaviour in both gas-free and
    gas-rich simulated discs (Bland-Hawthorn \& Tepper-Garcia 2021; Tepper-Garcia et al 2022).
\end{abstract}

\input{potential}
\newpage
\input{torque}
\newpage
\input{equations_of_motion}
\newpage
\input{experimental_setup}

\end{document}
